---
author:
- Arthur Dolgopolov
- Mikhail Freer
- Maggie Lehr
bibliography:
- 'refs.bib'
title: RSP$^2$
---

-   [Definitions](#definitions)
    -   [Altruistic Preferences](#altruistic-preferences)
    -   [Justice Seeking Preferences](#justice-seeking-preferences)
    -   [Inequality Averse Preferences](#inequality-averse-preferences)
-   [Results](#results)
    -   [Altruistic Preferences](#altruistic-preferences-1)
    -   [Justice Seeking Preferences](#justice-seeking-preferences-1)
    -   [Inequality Averse
        Preferences](#inequality-averse-preferences-1)

Definitions
===========

Consider an agent who must make a choice, $x\in {\mathbb{R}}^N_+$, to 
distribute resources between herself and $N$ other players. Let
$x = (x_s,x_O)$, where $x_s\in {\mathbb{R}}_+$ determines player’s own
payoff and $x_O \in {\mathbb{R}}^{N-1}_+$ is the payoff to all other
players. We assume that a player is choosing from the budget set (or the
set of available actions), denoted by $B$. Given the $\geq$ partial order of the 
downward closure of the budget constraint over ${\mathbb{R}}^N_+$, defined by
$$B_{\geq} = \{y: \text{ there is } x\in B \text{ such that } y \geq x\}$$
such that $B_{>}$ is the interior of the downward closure.

In practice we observe only finite set of choices. Dataset
$D=\{(x^t,B^t)\}$ is the collection of choices from available budgets. In the
general case we assume the budgets to be compact. In addition, if
we consider a uniformly distributed decision, in which every dollar is
equally distributed among all other players despite their initial endowments,
we only need the expenditures to be distributed at an equal rate.[^1] 
We refer to such experiment as to *uniform
experiment*. This is a common feature of the labor supply
subject to re-distributive taxation, or when choosing the optimal tax rates.

We are looking for the utility function which is consistent 
with a particular theory and rationalizes the observed choices made 
by the agent. This means that chosen point is at least as good as 
every choice in the budget set. Next, we
provide detailed definitions for every theory considered.

Altruistic Preferences
----------------------

In this case, a player gets utility from her own payoff as well as the payoff of
every other player. We explicitly refer to these preferences as
altruistic because we do not require, in general, the budget to be equal to
it’s downward closure. If it is the case, then the assumption of
monotonicity of the utility function can be reduced to the standard
notion of other-regarding preferences. That is, utility function is 
locally non-satiated.

\[def:OR\] A dataset $D=\{(x^t,B^t)\}$ is rationalizable with altruistic
preferences if there is a continuous and monotone utility function
$u:{\mathbb{R}}^N_+\rightarrow {\mathbb{R}}$ such that
$u(x^t) \in {\operatornamewithlimits{argmax}}\limits_{x\in B^t} u(x)$.

Justice Seeking Preferences
---------------------------

We assume that a person gets utility from her own payoff and the perceived 
justice in the society. We use a general concept of justice consistent with
the notion of justice is some social welfare function, which aggregates the payoffs
the entire group receives [@becker2013revealed]. This definition of the 
notion of justice we use allows for many different distributional preferences.
For instance, both utilitarian and egalitarian preferences can be 
classified as justice seekend preferences. 

Denote $S_n$ as the space of all
possible permutations of the vector of length $n$ and by $\sigma\in S_n$
a permutation.

\[def:NotionOfJustice\] A continuous function
$f:{\mathbb{R}}^{N-1}\rightarrow {\mathbb{R}}$ is said to be a justice
function if it satisfies

-   monotonicity: $x\geq (>) x'$ implies $f(x)\geq (>) f(x')$ for every
    $x,x'\in {\mathbb{R}}^{N-1}_+$, and

-   symmetric: $f(x)=f(\sigma(x))$ for every $\sigma\in S_{N-1}$.

Using this notion of justice, a choice can be consistent with justice seeking
preferences if it satisfies Pareto ordering and is anonymous. 
These general specification are satisfied by
most social welfare functions. However, let us note that the notion of
justice does not depend on the payoff the player receives herself. The 
interdependence between the distribution of payoffs in the
society and the payoff the player receives is done via the utility
function.[^2]

\[def:JS\] $f:{\mathbb{R}}^{N-1}\rightarrow {\mathbb{R}}$ be a justice
function. A dataset $D=\{(x^t,B^t)\}$ is rationalizable with justice
seeking preferences if there is a continuous and monotone utility
function $u(x_s,f(x_O))$ such that
$u(x^t) \in {\operatornamewithlimits{argmax}}\limits_{x\in B^t} u(x)$.

Inequality Averse Preferences
-----------------------------

Next, we consider a person who has notion of justice that is inequality averse. 
This means the agent would a transfer of resources from an agent with a 
larger endowment to an agent with a lower endowment, such that the 
ordering of people is preserved. This property is referred 
to in the literature on inequality measures as the
Pigou-Dalton (transfer) principle. We can define the inequality
measure as,

(THERE SHOULD BE CODE FOR A DEFINITION HERE)
A continuous function $f:{\mathbb{R}}^{N-1}\rightarrow {\mathbb{R}}$ is
said to be an inequality measure if it satisfies

-   inequality aversion: $x$ can be obtained (in relative income terms)
    from $x'$ by a sequence of Pigou-Dalton transfers implies
    $f(x)\geq (>) f(x')$ for every $x,x'\in {\mathbb{R}}^{N-1}_+$, and

-   symmetric: $f(x)=f(\sigma(x))$ for every $\sigma\in S_{N-1}$.

Note that inequality measures do not take into account the possible
dominance relations over the income distributions. Both
properties can be merged upon necessity (see Appendix X for more
details). However, this would deliver the theory which is weaker than
both inequality aversion and justice seeking preferences.

\[def:IA\] $f:{\mathbb{R}}^{N-1}\rightarrow {\mathbb{R}}$ be an
inequality measure. A dataset $D=\{(x^t,B^t)\}$ is rationalizable with
inequality averse preferences if there is a continuous and monotone
utility function $u(x_s,f(x_O))$ such that
$u(x^t) \in {\operatornamewithlimits{argmax}}\limits_{x\in B^t} u(x)$.

Results
=======

Next, we present the revealed preference tests for the three theories of
social preferences listed above. We start with presenting the test for
altruistic preferences, next we present the test for the justice seeking
preferences, and finally the test for the inequality averse preferences.

Altruistic Preferences
----------------------

Prior studies of altruistic preferences ([@afriat1967construction], 
[@nishimura2017unified]) have shown that rationalization with altruistic
preferences is equivalent to the Generalized Axiom of Revealed
Preferences [introduced by @varian1982nonparametric].

\[def:GARP\] A dataset satisfies GARP if for every sequence
$x^{t_1},\ldots,x^{t_n}$ such that $x^{t_j} \in B^{t_{j+1}}_{\geq}$ for
every $j\in \{1,\ldots,n-1\}$ then $x^{t_n} \notin B^{t_1}_{>_{t_1}}$.

\[[@afriat1967construction; @varian1982nonparametric]\] \[thm:AP\] A
dataset is rationalizable with altruistic preferences if and only if it
satisfes GARP.

Justice Seeking Preferences
---------------------------

\[def:SyGARP\] A dataset satisfies Sy-GARP if for every sequence
$x^{t_1},\ldots,x^{t_n}$ such that
$(x^{t_j}_s,\sigma(x_O^{t_j})) \in B^{t_{j+1}}_{\geq}$ for some
$\sigma\in S_{N-1}$ and for every $j\in \{1,\ldots,n-1\}$ then there is
no $\sigma\in S_{N-1}$ such that
$(x^{t_n}_s, \sigma(x^{t_n}_O)) \notin B^{t_1}_{>_{t_1}}$.

\[thm:JS\] If an experiment is rationalizable with justice seeking
preferences, then it satisfies Sy-GARP. Moreover, if data comes from the
uniform experiment, then the reverse is also true.

Inequality Averse Preferences
-----------------------------

consider Lorenz curve, denoted by $L(u)$, where $u\in[0,1]$ are the incomes in
normalized (by the maximum income) terms. We consider $L$ to
*dominate* $L'$ if $L(u)\geq L'(u)$ for every $u\in[0,1]$.

\[def:LDGARP\] A dataset satisfies LD-GARP if for every sequence
$x^{t_1},\ldots,x^{t_n}$ such that $x^{t_j}_s\geq x^{t_{j+1}}_s$ and
$L^{t_j}(u)$ dominates $L^{t_{j+1}}(u)$ for some $\sigma\in S_{N-1}$ and
for every $j\in \{1,\ldots,n-1\}$ then $x^{t_1}_s\leq x^{t_{n}}_s$
or/and $L^{t_1}(u)$ not dominates $L^{t_{n}}(u)$.

\[thm:IA\] If an experiment is rationalizable with inequality averse
preferences, then it satisfies LD-GARP. Moreover, if data comes from the
uniform experiment, then the reverse is also true.

[^1]: In more general setting we can allow the money to be distributed
    at constant pace for only sub-population. Moreover, it can be
    possible that money are distributed with different weights to the
    different parts of the population, however, the player herself can
    only decide on the total amount of money to be distributed, but not
    the detailed distribution to each member.

[^2]: If we include $x_s$ into the justice function, all the further
    results would still be necessary, being also sufficient for special
    cases. For more details on this case see [@becker2013revealed]
